---
title: "STAT 332 A1Q4"
author: "Yuqi Gu"
date: "6/2/2024"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
classurv <- read.csv("~/Desktop/STAT 332/Assignments/A1/classurv.txt", sep="")
library(survey)
```

### (a)
```{r}
classurv$GPA[classurv$GPA == -9] <- NA
N <- 350
svy <- svydesign(ids = ~1, probs = c(nrow(classurv)/N), fpc = rep(N, nrow(classurv)), 
                 data=classurv)
summary(svy)
GPA_mean <- svymean(~GPA, svy, na.rm = TRUE)
GPA_mean
dfall <- nrow(subset(classurv, !is.na(GPA)))-1
confint(GPA_mean, "GPA", df=dfall)
```
The point estimate for average GPA of all students is 3.006. The corresponding 95% CI is [2.90498, 3.107103].

### (b)
```{r}
male_students <- classurv[classurv$gender==1,]
n_male <- nrow(male_students)
male_svy <- svydesign(ids = ~1, probs = c(n_male/N), fpc=rep(N,nrow(male_students)), 
                     data=male_students)
male_mean <- svymean(~GPA, male_svy, na.rm = TRUE)
male_mean
dfmale <- nrow(subset(classurv[classurv$gender==1,], !is.na(GPA)))-1
confint(male_mean, "GPA", df=dfmale)
```
The point estimate for average GPA of all male students is 3.0491. The corresponding 95% CI is [2.908975, 3.189286].


### (c)
```{r}
female_students <- classurv[classurv$gender==2,]
n_female <- nrow(female_students)
female_svy <- svydesign(ids = ~1, probs = c(n_female/N), fpc=rep(N,nrow(female_students)), 
                      data=female_students)
female_mean <- svymean(~GPA, female_svy, na.rm = TRUE)
female_mean
dffemale <- nrow(subset(classurv[classurv$gender==2,], !is.na(GPA)))-1
confint(female_mean, "GPA", df=dffemale)
```
The point estimate for average GPA of all female students is 2.9664. The corresponding 95% CI is [2.801902, 3.130898].


### (d)
```{r}
spmale<-svyciprop(~I(gender==1), svy, method="mean", na.rm = TRUE)
spmale
spfemale<-svyciprop(~I(gender==2), svy, method="mean", na.rm = TRUE)
spfemale
```
The point estimate for the proportion of male students is 0.404. The corresponding 95% CI for the proportion of male students is [0.283, 0.52]. The point estimate for the proportion of female students is 0.596. The corresponding 95% CI for the proportion of female students is [0.476, 0.72].


### (e)
```{r}
svyyear1<-svyciprop(~I(year==1), svy, method ="logit", na.rm = TRUE)
svyyear1
svyyear2<-svyciprop(~I(year==2), svy, method ="logit", na.rm = TRUE)
svyyear2
svyyear3<-svyciprop(~I(year==3), svy, method ="mean", na.rm = TRUE)
svyyear3
svyyear4<-svyciprop(~I(year==4), svy, method ="mean", na.rm = TRUE)
svyyear4
svyyear5<-svyciprop(~I(year==5), svy, method ="logit", na.rm = TRUE)
svyyear5
```
The point estimates for the proportions of students that are freshman/sophomore/junior/senior/other are respectively 0.1404, 0.1228, 0.439, 0.263,0.03509.
The corresponding 95% CI for the proportions of students that are freshman/sophomore/junior/senior/other are respectively [0.0746, 0.25], [0.0622, 0.23], [0.317, 0.56], [0.155, 0.37], [0.00952, 0.12].
Note that since the point estimates for the proportions of students that are freshman/sophomore/other are smaller than 0.2, we use logit. 

### (f) 
```{r}
svyfre <- classurv[classurv$year==1,]
n_svyfreshman <- nrow(svyfre)
freshman_svy <- svydesign(ids = ~1, probs = c(n_svyfreshman/N), fpc=rep(N,nrow(svyfre)), 
                     data=svyfre)

svyciprop(~I(gender==1), freshman_svy, method ="logit", na.rm = TRUE)
svyciprop(~I(gender==2), freshman_svy, method ="logit", na.rm = TRUE)
```
Amongst freshmen, the point estimate for the proportions of male students is 0.125. The point estimate for the proportions of female students is 0.875. 
Since amongst freshmen, the point estimate for the proportion of male students is smaller than 0.2 and the point estimate for the proportion of female students is larger than 0.8, we use logit. The corresponding 95% CI for the proportion of male students is [0.00978, 0.67] and the corresponding 95% CI for the proportion of female students is [0.326, 0.99].

### (g) 
```{r}
svysen <- classurv[classurv$year==4,]
n_svysenior <- nrow(svysen)
senior_svy <- svydesign(ids = ~1, probs = c(n_svysenior/N), fpc=rep(N, nrow(svysen)), 
                     data=svysen)

svyciprop(~I(gender==1), senior_svy, method="mean", na.rm = TRUE)
svyciprop(~I(gender==2), senior_svy, method="mean", na.rm = TRUE)
```
Amongst seniors, the point estimate for the proportions of male students is 0.6. The corresponding 95% CI for the proportion of male students amongst senior is [0.325, 0.87]. The point estimate for the proportions of female students is 0.4. The corresponding 95% CI for the proportion of female students amongst seniors is [0.125, 0.67].

### (h) 
```{r}
topperformers <- classurv[classurv$GPA>=3.2,]
n_top <- nrow(topperformers)
top_svy <- svydesign(ids = ~1, probs = c(n_top/N), fpc=rep(N, nrow(topperformers)), 
                     data=topperformers)

svyciprop(~I(gender==1), top_svy, method="mean", na.rm = TRUE)
svyciprop(~I(gender==2), top_svy, method="mean", na.rm = TRUE)
```
Amongst the top permormers(i.e., GPA >= 3.2), the point estimate for the proportions of male students is 0.5. The corresponding 95% CI for the proportion of male students amongst senior is [0.263, 0.74]. The point estimate for the proportions of female students is 0.5. The corresponding 95% CI for the proportion of female students amongst seniors is [0.263, 0.74].