---
title: "Q4"
author: "Yuqi Gu"
date: "7/13/2024"
output:
  html_document: default
  pdf_document: default
header-includes: usepackage{multirow}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### (a)
```{r}
col1 <- c(0,0,0,0,0,0,0,1,1,1,1,1,1,1,1,1,1,2,2,2,2,2,2,2,2,2,2,3,3,3,3,3)
col2 <- c(38.14, 38.2, 38.31, 38.14, 38.29, 38.17, 38.2,
               38.28, 38.17, 38.08, 38.25, 38.18, 38.03, 37.95, 38.26, 38.3, 38.21,
               38.17, 38.13, 38.16, 38.3, 38.34, 38.34, 38.17, 38.18, 38.09, 38.06,
               38.14, 38.3, 38.21, 38.04, 38.37)
but.dat <- data.frame(pushes_num = col1, wait_time=col2)
but.dat$pushes_num <- as.factor(but.dat$pushes_num)
but.aov <- aov(wait_time ~ pushes_num, data = but.dat)
summary(but.aov)
```
Thus, the ANOVA table is:
Source    d.f.    SS          MS
Treatment 3       0.00805     0.002682
Error     28      0.30595     0.010927 
Total     31      0.314


### (b)
```{r}
qf(0.95, 3, 28)
```



### (c)
```{r}
tmp <- model.tables(but.aov, type="means")
but.tab <- data.frame(names(tmp$tables[[2]]), c(tmp$tables[[2]]), tmp$n[[1]])
colnames(but.tab) <- c(names(tmp$tables[2]), "means", "ni")

but.tab$var <- sum(but.aov$res^2)/but.aov$df.residual*1/c(tmp$n[[1]])
but.tab$sd <- sqrt(but.tab$var)

but.tab$lower95 <- but.tab$means-qt(0.975, but.aov$df.residual) * but.tab$sd
but.tab$upper95 <- but.tab$means+qt(0.975, but.aov$df.residual) * but.tab$sd
but.tab
```
Thus, the mean waiting time after pushing the button once is 38.171. The corresponding 95% CI for this estimate is [38.10329, 38.23871]. The mean waiting time after pushing the button three times is 38.212. The corresponding 95% CI for this estimate is [38.11624, 38.30776].

### (d)
```{r}
TukeyHSD(but.aov)
```
Based on the output, we find that the pairwise comparison between the two mean waiting times of part (c) is 0.041, the number in the fifth row of the first column called diff. The corresponding 95% CI is [-0.1153224, 0.1973224]. Since the 95% CI contains 0, we conclude that the number of pushes of 1 or 3 of the pedestrian button has no effect on the waiting time for the walk signal.

### (e)
```{r}
MSE <- 0.010927
L_hat <- (but.tab$means[1] - (but.tab$means[2]+but.tab$means[3]+but.tab$means[4])/3)
var <- MSE*(1/7+(-1/3)^2/10+(-1/3)^2/10+(-1/3)^2/5)
t <- qt(0.975, 28)
lwr <- L_hat - t* sqrt(var)
upr <- L_hat + t* sqrt(var)
L_hat
lwr
upr
```

The estimate of contrast is 0.01480952. The corresponding 95% CI is [-0.07786012, 0.1074792]. Since this CI contains 0, we conclude that the effect of no pushes or pushing the button once on the waiting time for the walk signal has no difference.


